The Conjunctive Disjunctive Graph Node Kernel for Disease Gene Prioritization

### 2017-11-16 neurocomputingv3.tex
corrected file

\subsection{Model Selection}
The hyper-parameters of the various methods were tuned using a 10-fold cross validation. However, due to the non i.i.d. nature of the problem, we emploied a stronger setup to ensure no information leakage. The dataset $0$ (Connective) on which we  validated the performance was never subsequently used in the predictive performance estimation. The values for the diffusion parameter $\beta$ in LEDK and MEDK were sampled in $\lbrace 10^{-3}, 10^{-3}, 10^{-2}, 10^{-1} \rbrace$, time steps $t$ in MDK in $\lbrace 1, 10, 100 \rbrace$ and RLK parameter $\beta$ in $\lbrace 1, 4, 7 \rbrace$. For CDNK, the degree threshold values, $D$, were sampled in $\lbrace 10,\ 15,\ 20 \rbrace$, clique size threshold, $C$, in $\lbrace 4,\ 5 \rbrace$, maximum radius, $r^*$, in $\lbrace 1,\ 2 \rbrace$, maximum distance, $d^*$, in $\lbrace 1, 2,\ 3,\ 4 \rbrace$, number of clusters $P$ in $\lbrace 5,\ 7 \rbrace$. Finally, the regularization trade off parameter in the SVM was sampled in $\lbrace 10^{-5},  \ 10^{-4}, \ 10^{-3},\ 10^{-2},\ 10^{-1}, 1,\ 10,\ 10^2 \rbrace$. The optimal parameter values of CDNK obtained by validation process are $D = 1$, $C=5$, $r^*=2$, $d^*=1$, $P=15$ and the trade-off parameter of SVM is $10^{-2}$.
%\subsection{Parameter Impact}
%In order to evaluate the stability of kernel w.r.t the change of each parameter values, we perform experiments with disease-gene association 2 on the BioGPS network. For each parameter, first we fix the values of all other parameters, but allow the value of considered papameter to be changed. More precisely, we use the optimal tuple $C = 5,\ d^* = 1,\ D = 20,\ P = 15$ for the fixed of parameters' values. We let the changes of $C$ in $\lbrace 4,\ 5,\ 6,\ 7,\ 8,\ 9\rbrace$, $d^*$ in $\lbrace 1,\ 2,\ 3,\ 4\rbrace$, $D$ in $\lbrace 8,\ 10,\  15,\ 20,\ 30,\ 40,\ 50\rbrace$ and real vector size, $P$ in $\lbrace 8,\ 10,\ 15,\ 20,\ 30,\ 50,\ 70, 90\rbrace$.

@@

Let now turn our attention to question {\it Q2}, i.e. whether the possibility to use local topology and/or real valued information is useful to improve performance.
The first part of the question, i.e. usefulness of local topology, can be answered by looking at results of CDNK3, where only discrete labels defined by the baseline approach are used. If we look at results obtained on BioGPS network, we see that CDNK3 outperforms diffusion-based kernels with the 
 exceptions of datasets $3$, $7$, $8$, and $9$, thus $7$ times over $11$. It is important to stress, however, that on the average CDNK3 outperformed the diffusion-based kernels of $10.19\%$, while it was outperformed on the average of $6.05\%$. An even better scenario is observed for the Pathways network, where only in one case (dataset $4$) the performance of CDNK3 was inferior of almost all diffusion-based kernels.
Thus, it seems that exploiting local topological information is actually important in many cases. 
Regarding the integration of real valued vectors, we have to look at variations CDNK2 and CDNK4, that exploit real valued labels based on GO.
It is readily evident that CDNK2 and CDNK4 improve on the performance of diffusion-based kernels in most cases. Specifically, for BioGPS network,
CDNK2 is the best performer in absolute in $7$ out of $11$ datasets. Moreover, its worst performance places it above all diffusion-based kernels.
In turn, CDNK4 is first in $3$ datasets and second in $4$ datasets, and again its worst performance places it above all diffusion-based kernels.
When looking at the results for the Pathways network, CDNK4 is the best performer among all kernels, while CNDK2 places in second position since
even if it is first on two datasets as well as RLK, RLK has a worst ranking score on most of the other datasets.
We can conclude that also the exploitation of real valued information seems to improve the performance. Thus, also for question  {\it Q2} we can give a positive answer.
%In particular, considering the use of discrete labels based on ontology, using side information helps to increase the performance in all diseases on BioGPS and in 9 out of 11 diseases on Pathways. Similarly, by employing auxiliary information associated to graph nodes, the performance of CNDK in case of using uniform discrete labels is also improved in 9 and 8 out of 11 on BioGPS and Pathways, respectively.

>>>>

\subsection{Model Selection}
The hyper-parameters of the various methods were tuned using a 10-fold cross validation. However, due to the non i.i.d. nature of the problem, we emploied a stronger setup to ensure no information leakage. The dataset $0$ (Connective) on which we  validated the performance was never subsequently used in the predictive performance estimation. The values for the diffusion parameter $\beta$ in LEDK and MEDK were sampled in $\lbrace 10^{-3}, 10^{-3}, 10^{-2}, 10^{-1} \rbrace$, time steps $t$ in MDK in $\lbrace 1, 10, 100 \rbrace$ and RLK parameter $\beta$ in $\lbrace 1, 4, 7 \rbrace$. For CDNK, the degree threshold values, $D$, were sampled in $\lbrace 10,\ 15,\ 20 \rbrace$, clique size threshold, $C$, in $\lbrace 4,\ 5 \rbrace$, maximum radius, $r^*$, in $\lbrace 1,\ 2 \rbrace$, maximum distance, $d^*$, in $\lbrace 1, 2,\ 3,\ 4 \rbrace$, number of clusters $P$ in $\lbrace 5,\ 7 \rbrace$. Finally, the regularization trade off parameter in the SVM was sampled in $\lbrace 10^{-5},  \ 10^{-4}, \ 10^{-3},\ 10^{-2},\ 10^{-1}, 1,\ 10,\ 10^2 \rbrace$. The optimal parameter values of CDNK obtained by validation process are $D = 20$, $C=5$, $r^*=2$, $d^*=1$, $P=15$ and the trade-off parameter of SVM is $10^{-2}$.
%\subsection{Parameter Impact}
%In order to evaluate the stability of kernel w.r.t the change of each parameter values, we perform experiments with disease-gene association 2 on the BioGPS network. For each parameter, first we fix the values of all other parameters, but allow the value of considered papameter to be changed. More precisely, we use the optimal tuple $C = 5,\ d^* = 1,\ D = 20,\ P = 15$ for the fixed of parameters' values. We let the changes of $C$ in $\lbrace 4,\ 5,\ 6,\ 7,\ 8,\ 9\rbrace$, $d^*$ in $\lbrace 1,\ 2,\ 3,\ 4\rbrace$, $D$ in $\lbrace 8,\ 10,\  15,\ 20,\ 30,\ 40,\ 50\rbrace$ and real vector size, $P$ in $\lbrace 8,\ 10,\ 15,\ 20,\ 30,\ 50,\ 70, 90\rbrace$.

@@

Let now turn our attention to question {\it Q2}, i.e. whether the possibility to use local topology and/or real valued information is useful to improve performance.
The first part of the question, i.e. usefulness of local topology, can be answered by looking at results of CDNK3, where only discrete labels defined by the baseline approach are used. If we look at results obtained on BioGPS network, we see that CDNK3 outperforms diffusion-based kernels with the 
 exceptions of datasets $3$, $7$, $8$, and $9$, thus $7$ times over $11$. It is important to stress, however, that on the average CDNK3 outperformed the diffusion-based kernels of $10.19\%$, while it was outperformed on the average of $6.05\%$. An even better scenario is observed for the Pathways network, where only in one case (disease gene association $4$) the performance of CDNK3 was inferior of almost all diffusion-based kernels.
Thus, it seems that exploiting local topological information is actually important in many cases. 
Regarding the integration of real valued vectors, we have to look at variations CDNK2 and CDNK4, that exploit real valued labels based on GO.
It is readily evident that CDNK2 and CDNK4 improve on the performance of diffusion-based kernels in most cases. Specifically, for BioGPS network,
CDNK2 is the best performer in absolute in $7$ out of $11$ datasets. Moreover, its worst performance places it above all diffusion-based kernels.
In turn, CDNK4 is first in $3$ datasets and second in $4$ datasets, and again its worst performance places it above all diffusion-based kernels.
When looking at the results for the Pathways network, CDNK4 is the best performer among all kernels, while CNDK2 places in second position since
even if it is first on two datasets as well as RLK, RLK has a worst ranking score on most of the other datasets.
We can conclude that also the exploitation of real valued information seems to improve the performance. Thus, also for question  {\it Q2} we can give a positive answer.
%In particular, considering the use of discrete labels based on ontology, using side information helps to increase the performance in all diseases on BioGPS and in 9 out of 11 diseases on Pathways. Similarly, by employing auxiliary information associated to graph nodes, the performance of CNDK in case of using uniform discrete labels is also improved in 9 and 8 out of 11 on BioGPS and Pathways, respectively.
